---
title: "presidential_money"
author: "martin-martin"
date: "29 Februar 2016"
output: html_document
---

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# loading the packages I use
library(ggplot2)
library(reshape2)
```

# Dollars for Votes - the US presidential elections

by Martin

========================================================

I'm currently spending my time in the USA, and it's (pre-)election time! Having a memebership to a 24h gym gets me exposed to a massive amount of politics on TV while I'm doing my time on the crosstrainer or even when I'm just changing in the dressing room.
It's February, the elections are in November, but there is a near-constant broadcast of candidates battling each other on CNN.

I'm impressed and surprised about this omnipresence of the elections on TV, already so early on. And I'm concerned about the rudeness that these debates often display. Politics in Europe are different. They're not golden, nope. And there are insults and dirty tricks played also there. But here in the US there's a level of "attack", "aggression" and "god" in the talks, that I am thoroughly not used to. :/

One of the times while in the dressing room, I got stuck watching the TV talk about **Dollars per Vote**, and found this comparison very interesting. Some candidates spend very little money as compared to others in absolute, and yet in the end some end up spending still more money _per vote_ received.

Now, my analysis will not deal with this comparison, since the ballots have not been cast yet (for a few months...), but I will make a preliminary analysis of the money that the presidential candidates were receiving from individual contributors in the run-up for the presidency.

## Pt. 0: The Data

The data for the 2016 Presidential Campaign Contributions can be found here: http://fec.gov/disclosurep/PDownload.do
Since I am currently in California, I chose and downloaded the dataset for `CA.zip`.

Here's also a peek at the structure of the data that the .csv contains. It's an excerpt from this file: ftp://ftp.fec.gov/FEC/Presidential_Map/2016/DATA_DICTIONARIES/CONTRIBUTOR_FORMAT.txt
I have also indicated which of the columns I ended up dropping.

```
The text file is comma delimited and uses double-quotation marks as the text qualifier.

------------------------------------------------------------------------ 

CMTE_ID			        COMMITTEE ID				                      S <- dropped
CAND_ID			        CANDIDATE ID				                      S
CAND_NM			        CANDIDATE NAME				                    S
CONTBR_NM		        CONTRIBUTOR NAME			                    S
CONTBR_CITY		      CONTRIBUTOR CITY			                    S
CONTBR_ST		        CONTRIBUTOR STATE			                    S <- dropped
CONTBR_ZIP		      CONTRIBUTOR ZIP CODE			                S
CONTBR_EMPLOYER		  CONTRIBUTOR EMPLOYER			                S
CONTBR_OCCUPATION	  CONTRIBUTOR OCCUPATION			              S
CONTB_RECEIPT_AMT	  CONTRIBUTION RECEIPT AMOUNT		            N
CONTB_RECEIPT_DT	  CONTRIBUTION RECEIPT DATE		              D	
RECEIPT_DESC		    RECEIPT DESCRIPTION			                  S <- dropped
MEMO_CD			        MEMO CODE				                          S <- dropped
MEMO_TEXT		        MEMO TEXT				                          S <- dropped
FORM_TP			        FORM TYPE				                          S <- dropped
FILE_NUM		        FILE NUMBER				                        N <- dropped
TRAN_ID			        TRANSACTION ID				                    S <- dropped
ELECTION_TP		      ELECTION TYPE/PRIMARY GENERAL INDICATOR   S <- dropped


Data Type:  S = string (alpha or alpha-numeric); D = date; N = numeric  

------------------------------------------------------------------------ 
```

One thing that becomes very obvious here, and that strongly influenced my exploration, is that there is **only one column containing numerical variables**: `contb_receipt_amt`.

This is the column that holds the amounts of the contributions in $, and it is therefore also one of the most interesting columns. A lot of comparisons were anchored to it.

However, there is also another "hidden" numerical value associated with the dataset, and that is the **number of rows**, that allows many graphs to be created in combination with the categorical values of the other columns.

### Ideas and Interests that I will explore

+ Who got the most contributions? Which person, which party, which gender?

- Which city contributed the most, and to whom (also: in relation to the number of citizens)

- How many "NOT EMPLOYED" people contributed to the campaigns as compared to people with employment; Also: whom did the "NOT EMPLOYED" support the most?

## Pt. 1: Mini-Wrangling

In order to be able to load the data, I had to add a comma at the end of the header row in my .csv file.

```{r echo=FALSE, Load_the_Data}
# Load the Data
setwd('~/Desktop/p4_US_campaign_funding')
df <- read.csv('P00000001-CA.csv')
names(df)
```

### Adding new columns: cand_gender and cand_party

For some of my questions, I am interested to take a look at the data also depending on the gender and the party affiliation of the different candidates. Since these columns are not present in the dataset, I went to gather the required information online and added it as two new columns to my data.frame `df`.

```{r echo=FALSE, Adding_Columns}
## PART1 - Mapping
# First I'll be mapping Candidate names to Candidate IDs. This I'll do so that I will be able to refer to a cand_id when passing a function, which is easier than writing out the full name.

# select only rows that are not duplicates in the cand_id value
candidates <- df[!duplicated(df$cand_id), ]
# define which columns to keep
keeps <- c("cand_id", "cand_nm")
# assigning the new mapping data.frame to a variable
cand_id_map <- candidates[keeps]
#cand_id_map

# how many candidates are in the primaries?
#nrow(cand_id_map)

# I'd also like to have the party associated with the candidate
democrats <- c('P60007168', 'P00003392', 'P60009685', 'P60008885', 'P60007671')
republicans <- c('P60006723', 'P60008521', 'P60008059', 'P60005915', 'P60003670', 'P80001571', 'P60007242', 'P60006111', 'P40003576', 'P20002721', 'P60008398', 'P80003478', 'P60007572', 'P60007697', 'P20003281', 'P60006046')
greens <- c('P20003984')

# double checking
#all_cands <- c(democrats, republicans, greens)
#duplicated(all_cands)
#nrow(cand_id_map) == length(democrats) + length(republicans) + length(greens)


## PART2 - Adding the Columns
# so now I can add this new column to the df.
# there is one member of the green party 'G', the republicans will get the value 'R'
# and the democrats the value 'D'
df$cand_party <- ifelse(df$cand_nm != 'Stein, Jill', ifelse(df$cand_id %in% republicans, "R", "D"), "G")
#head(subset(df, cand_party == "G"))

# and also the gender of the candidate
female <- c('Fiorina, Carly', 'Clinton, Hillary Rodham', 'Stein, Jill')
# adding a new column to the df
df$cand_gender <- ifelse(df$cand_nm %in% female, "F", "M")

# checking it out
#head(subset(df, cand_nm == 'Fiorina, Carly'))
#names(df)
```

### Removing obsolete columns

There is quite a big amount of data that I am not intending to use for my exploration. These include e.g. `conbr_st`, which is in my case always `CA`, or `file_num`, a unique number used to link the transactions to the reports (which I'll also drop).

```{r echo=FALSE, Removing_Columns}
discard <- c('cmte_id', 'contbr_st', 'receipt_desc', 'memo_cd', 'memo_text', 'form_tp', 'file_num', 'election_tp', 'X')
df <- df[ , !(names(df) %in% discard)]
#names(df)
```

## Pt. 2: Exploration

After having cleaned and adapted the dataset to my wishes, I can start into my exploration.

For example, I can query how many individual contributions did each candidate receive:

```{r echo=FALSE, First_Steps}
# E.g. for Hillary Clinton:
sum(df$cand_id == 'P00003392')
```

### QUESTION 1: _Who got the most money through contributions?_

In this section I'll take a look at money and amounts. I'm wondering which candidate got **the most contributions**, and is it the same person who got **the most money** through contributions? How is the size of the contributions distributed in regards to the candidates?
Which gender received more contributions, or more money? Which political party?

```{r echo=FALSE, Exploring_Most_Money}
### PART 1: the mean and distributions

# calculating the mean of the contributions for each candidate
mean_contrib_per_cand <- aggregate(contb_receipt_amt ~ cand_nm, data = df, mean)
# adding a count of contributions
mean_contrib_per_cand$count <- table(df$cand_nm)
#mean_contrib_per_cand

# cool graph showing the distributions!
ggplot(df, aes(cand_nm, contb_receipt_amt)) +
  geom_boxplot()

# doing a frequency plot makes me see something about Sanders (big dot low down), that some names have a small distribution, and some a wider one, another person with big dot at the bottom, yet also a quite big one higher up (i think this is Clinton); and one person with a very high distribution up and down, but also a strong base it seems.
ggplot(df, aes(contb_receipt_amt, cand_nm)) +
  geom_count()

# this graph could be interesting
ggplot(mean_contrib_per_cand, aes(cand_nm, contb_receipt_amt)) +
  geom_density() + 
  geom_label(aes(label=cand_nm))

########################## TAKE THIS ONE ##########################
# this is how it works, no need for the labels (cleaned it up a bit)
# and reordered ascending! this plot gives a nice overview of the mean contributions per candidate! maybe it's a keeper ; )
ggplot(mean_contrib_per_cand, aes(reorder(cand_nm, -contb_receipt_amt), contb_receipt_amt)) +
  geom_density() + 
  coord_flip()
# label is a cool way to do it, but still!



### PART 2: amounts

# most of the contributions are below 1000 $
ggplot(df, aes(contb_receipt_amt)) +
  geom_freqpoly()

# trying around some aggregate()
# sum all receipt amounts, aggregating rows based on candidate name
money_by_cand <- aggregate(contb_receipt_amt ~ cand_nm, data = df, sum)
# nice this works, so now we can make a graph, maybe interesting with color and alpha
ggplot(aes(x = contb_receipt_amt, y = cand_nm, alpha=0.1, color=cand_nm), data = df) +
  geom_jitter() +
  scale_x_continuous() +
  # removing the legends for better visibility
  guides(color=FALSE, alpha=FALSE)


#### PART 2.1: refunds

# this gives me negative values for amounts, which is very confusing.
neg_contb_amt <- subset(df, contb_receipt_amt < 0)
#head(neg_contb_amt)
# sum all receipt amounts, aggregating rows based on candidate name
only_refunds <- subset(df, contb_receipt_amt < 0)
refunds_by_contb <- aggregate(contb_receipt_amt ~ contbr_nm, data = only_refunds, sum)
#head(refunds_by_contb)
nrow(refunds_by_contb)
# seems there are nearly 1500 "contributors" who actually got more money refunded than what they gave.
nrow(only_refunds)
# and some gave some, but also got some back (or got some back multiple times?)

# tried a plot for these refunds (which was too big, so I'll take a look at the most $$)
ggplot(subset(refunds_by_contb, contb_receipt_amt < -10000), aes(contb_receipt_amt, contbr_nm)) + geom_point()
# this is a lot of money to get back. I wonder why and how. But I must admit I don't really understand these US politics and fundings of the candidates
# Here's some: https://ballotpedia.org/California_Proposition_34,_Limits_on_Campaign_Contributions_(2000)


#### PART 2.2: contributions

# Well, so let's see who *gave* the most
only_contribs <- subset(df, contb_receipt_amt > 0)
ggplot(subset(only_contribs, contb_receipt_amt > 10000), aes(contb_receipt_amt, contbr_nm)) + geom_point()
top_givers <- subset(only_contribs, contb_receipt_amt > 10000)
nrow(top_givers)
# so there are 24 people that all gave the same amount, so I suspect that there is an upper limit around 10.000$ (maybe without taxes, or such? They should be alread PACs?)

# I also spotted potentially erroneous data, with the contributor DE GROOTE, DOUG MR. being listed, as well as a DE GROOTE, DOUG, both with the same amount, which makes me believe that it is a mistake in the data (or a rather badly executed way of increasing one's contribution limit...)
# maybe can check :)
#top_givers
duplicated(top_givers$tran_id)
# Yep, seems that some of those transactions are listed more often in the dataset.
# actually quite a lot of them! 7 TRUE!
dup_givers <- subset(top_givers, duplicated(top_givers$tran_id) == TRUE)
dup_givers
# Seems that DE GROOTE could be also a company, because it's listed also among the contbr_employer column.
# Well. But I'm not gonna go hunt down these individuals :)



### PART 3: gender and money

# number of contributions per gender of candidate
ggplot(df, aes(cand_gender)) + geom_bar() + scale_x_discrete()
# total amount of money through contributions per gender of candidate
gender_money <- aggregate(contb_receipt_amt ~ cand_gender, data = df, sum)
#head(gender_money)
ggplot(gender_money, aes(cand_gender, contb_receipt_amt)) + 
  geom_bar(stat='identity', fill = c('pink', 'lightblue')) + scale_x_discrete()
# whoa! seems that females received nearly as much money in contributions as males did, even though there are only 3 female vs. 19 male candidates + there were way less fewer contributions to femal candidates than to male candidates


#### PART 3.1: binning

# so it could be interesting to take a look at the mean contribution per F/M cand
summary(subset(df, cand_gender == 'F')$contb_receipt_amt)
summary(subset(df, cand_gender == 'M')$contb_receipt_amt)

summary(df$contb_receipt_amt)
# min = -10000 and max = 10800
# i'll start with a generous binwidth of 1000
seq(0, 11000, by = 1000)
# making bins for all the money
df$money_bin <- cut(df$contb_receipt_amt, breaks = seq(0, 11000, by = 1000))
head(df)
# we can make this look nicer using a function:
money.cat <- function(x, lower = 0, upper, by = 1000,
                   sep = "-", above.char = "+") {

 labs <- c(paste(seq(lower, upper - by, by = by),
                 seq(lower + by - 1, upper - 1, by = by),
                 sep = sep),
           paste(upper, above.char, sep = ""))

 cut(floor(x), breaks = c(seq(lower, upper, by = by), Inf),
     right = FALSE, labels = labs)
}

# and assign the values with the prettified labels to the new colum
df$money_bin <- money.cat(df$contb_receipt_amt, upper = 11000)
head(df)
# looks better now :)

ggplot(df, aes(money_bin)) + geom_bar(stat = 'count')
# so this graphic is pretty useless - nearly every contribution falls within the first bin, there are very few that are above 1000.
# so i can adjust the binnning:
df$money_bin <- cut(df$contb_receipt_amt, breaks = c(0, 100, 200, 300, 400, 500, 600, 700, 800, 900, 1000, 2000, 3000, 11000))
#head(df)
ggplot(df, aes(money_bin)) + geom_bar(stat = 'count')
# whoa! most contributions are actually between 0-100$!!
# So there are many many samll contributions that were made in CA
table(df$money_bin)
# Interesting, because I already saw that most low contributions went to one candidate: Sanders, Bernard.


#### PART 3.2: proportions

# Display proportions of female/male candidates in the respective parties
ggplot(df, aes(x = cand_gender, y = cand_party)) + 
  geom_count(aes(size = ..prop.., group = cand_party)) +
  scale_size_area(max_size = 10)
# Here the proportions on how many candidates of the respective gender are running in the presidential elections is very different between the three parties.

# let's see a graph plotting the statistical percentage of money per candidate per gender
gender_money$gen_mon_perc <- ifelse(gender_money$cand_gender == 'F', gender_money$contb_receipt_amt / 3, gender_money$contb_receipt_amt / 19)
gender_money
ggplot(gender_money, aes(cand_gender, contb_receipt_amt)) +
  scale_x_discrete() +
  geom_bar(stat='identity', alpha=0.5) +
  geom_bar(aes(cand_gender, gen_mon_perc), stat='identity', fill=c('pink', 'lightblue'), alpha=0.5)

# but of course that's VERY misleading!... Clinton got so much of the amounts contributed to female candidates, that we should make this clearer:
gender_cand_money <- aggregate(contb_receipt_amt ~ cand_gender + cand_nm, data = df, sum)
#gender_cand_money
ggplot(gender_cand_money) + scale_x_discrete() +
  geom_bar(aes(cand_gender, contb_receipt_amt, fill=cand_nm), stat='identity', alpha=0.8)
# whoa! as we can see, Hillary Clinton received by far the largest amount of money through contributions. she nearly single-handedly takes half of all the contributions made



### PART 4: party money

########################## TAKE THIS ONE ##########################
# I can also see, that Clinton and Sanders make the two biggest single sections, and they are both with the democrats. therefore it would be interesting to plot money per party instead of gender
party_cand_money <- aggregate(contb_receipt_amt ~ cand_party + cand_nm, data = df, sum)
ggplot(party_cand_money) + scale_x_discrete() +
  geom_bar(aes(cand_party, contb_receipt_amt, fill=cand_nm), stat='identity', alpha=0.8)

# so we can see that the democrats received way more money through contributitons than the republicans, and Hillary Clinton alone received more money than all the republican candidates combined.
# whereas the contributions to the green party are so tiny, that they become invisible in this visualization

# TODO: could try to clean this up a bit, by keep only the more-earner-candidates



### PART 5: candidate money

# number of contributions per candidate
qplot(cand_nm, data = subset(df, table(df$cand_nm) > 500))

# how many contributions per candidate
table(df$cand_nm)
ggplot(df, aes(contb_receipt_amt)) +
  geom_histogram() +
  facet_wrap(~cand_nm)
# while most are vanishingly small in comparison, four candidates stick out in terms of number of contributions

ggplot(df, aes(sum(contb_receipt_amt))) +
  geom_histogram() +
  facet_wrap(~cand_nm)
# here's the amounts; not very telling these plots.

```

### QUESTION 2: _Which city contributed the most, and to whom (and in relation to the inhabitants, maybe)_

```{r}
ggplot(df, aes(contb_receipt_amt, contbr_city)) + scale_y_discrete() +
  geom_bin2d()

# this one is actually quite interesting:
ggplot(df, aes(contb_receipt_amt, contbr_city)) + scale_y_discrete() +
  geom_count()
# shows how people give distinct (full numbers) amounts of money often (with the clear lines down)


########################## TAKE THIS ONE ##########################
# jitter does a similar thing as geom_point (or geom_count), here i'll apply it only to the actual contributions, leaving out the refunds
# and set an alpha to see better the density of the distribution
ggplot(only_contribs, aes(contb_receipt_amt, contbr_city)) + scale_y_discrete() +
  geom_jitter(alpha=0.1)
# this is a nice graph :)
# it nicely tells something about how much do most people donate (first big bar right of 0), about which cities had the most contributions (horizontal black lines), and where are there common discrete jumps, maybe related to regulations, such as donation limits (vertical lines)

# TODO: clean it up, remove cities (keep only the ones that gave the most money in total), research regarding contribution limits; maybe color dots regarding which party was donated to

all_cities <- aggregate(contb_receipt_amt ~ contbr_city, data=only_contribs, sum)
rich_cities <- subset(all_cities, contb_receipt_amt > 150000)
nrow(rich_cities)
head(rich_cities)

ggplot(rich_cities, aes(contb_receipt_amt, contbr_city)) + scale_y_discrete() +
  geom_point()
# maybe this is not so interesting, and i should rather right away put it in relation with the number of inhabitants. but that's a lot of work to gather all this data...


### Getting Population estimates for CA

# found some here: https://www.census.gov/popest/data/cities/totals/2012/SUB-EST2012.html
# the data is for 2012, but it's the most current one listing *cities* that I found there.
# It's not perfect, but I just want to take a look :)

pop_df <- read.csv('population_CA_2012.csv')
names(pop_df)
head(pop_df)
# i'm only gonna deal with NAME and the latest estimate
pop_df <- pop_df[, c('NAME', 'POPESTIMATE2012')]
pop_df
# yay to stackoverflow! and the internet in general! :)
# removing the last word ('city') from the city names
pop_df$NAME <- sub(pop_df$NAME, pattern = " [[:alpha:]]*$", replacement = "")
# converting the strings in the column to uppercase to match how they are stored in my df
pop_df$NAME <- toupper(pop_df$NAME)
# merge the two data.frames by the city-name columns
only_contribs <- merge(only_contribs, pop_df, by.x = 'contbr_city', by.y = 'NAME')
#head(only_contribs)
head(only_contribs)

ggplot(only_contribs, aes(POPESTIMATE2012, contb_receipt_amt)) + geom_bin2d()
# would be interesting to display this with a facet_wrap. however, i believe there are too many cities. so i should get the ratios first? boh, bring it.


### this would totally be a multivariate analysis... just gotta manage it!
?geom
?facet_wrap
# this would be cool maybe, but it's simply too many cities...
ggplot(only_contribs, aes(POPESTIMATE2012, contb_receipt_amt)) + 
  geom_count() +
  facet_wrap(~ contbr_city)
#  geom_count(aes(size = ..prop.., group = cand_party)) +
#  scale_size_area(max_size = 10)

# getting the ratios
total_contribs <- aggregate(data=only_contribs, contb_receipt_amt ~ contbr_city, sum)
total_contribs <- merge(total_contribs, pop_df, by.x = 'contbr_city', by.y = 'NAME')
head(total_contribs)
total_contribs <- total_contribs[!duplicated(total_contribs), ]
head(total_contribs)
# oops, there is also data for "counties" in here, that have now the same name as some of the cities... should remove the rows with the higher data.


ggplot(only_contribs, aes(POPESTIMATE2012, contbr_city)) + geom_density()

```

### Question 3: How much and to whom did NOT EMPLOYED give

I am very perplexed and involved with the topic of homeless people in the US, so anything that goes into that direction rings a bell with me.
Here I'm trying to investigate a little bit into the political direction that homeless might be having a tendency for.
However, I understand that this is highly hypothetical, because I only have data of **monetary contributions**, that **"NOT EMPLOYED"** people gave for the presidential campaigns.
Of course giving money != political orientation (It might be a good proxy, however what I'm trying to say is, that it's not an exhaustive factor. Many people might have a political orientation, however did not contribute to the campaigns monetarily. This might be especially true for homeless people, who are very likely to have very little money at their hands).
Further, NOT EMPLOYED != homeless. There are quite a few people that are employed, but homeless in the US. Assuming that they would give a contribution, they would fall into a different category.

These graphs are not gonna show much about how much money was flowing, but rather is intended as a proxy on where does a certain section of society lean towards politically.


```{r}
# adding a column containing a pseudo-boolean gleaned from the employment data
df$contb_emp_status <- ifelse(df$contbr_occupation == 'NOT EMPLOYED', 'not employed', 'employed')
#head(df)
# aggregating the amounts of money for each of the two groups
empl_money <- aggregate(contb_receipt_amt ~ contb_emp_status, data=df, sum)

# who gave how much?
ggplot(empl_money, aes(contb_emp_status, contb_receipt_amt)) + scale_x_discrete() +
  geom_bar(fill=c('gray', 'lightblue'), stat = 'identity')
# okay, not so exciting :)
# wait
table(df$contb_emp_status)
# there are very few NOT EMPLOYED compared to those with employment, so I'd need ratios

# Display proportions of which party did people with/without employment give contributions to
d <- ggplot(df, aes(x = contb_emp_status, y = cand_party))
d + geom_count(aes(size = ..prop.., group = cand_party, color=cand_party)) +
  scale_size_area(max_size = 80)
# interesting: seems that a much higher percentage of people without employment contribute to the democrats
# NOTE: I boosted the max_size variable, which makes it even more clear how vanishingly small is the percentage of NOT EMPLOYED people that contributed to the Republican party.



########################## TAKE THIS ONE ##########################
# and well, let's see for whom...
d <- ggplot(df, aes(x = contb_emp_status, y = cand_nm))
d + geom_count(aes(size = ..prop.., group = cand_party, color=cand_party)) +
  scale_size_area(max_size = 10)
# bam! Sanders, Bernard gets percentage-wise the most contributions from the NOT EMPLOYED!
# and it seems that apart from Trump, Donald J. there is no Republican left in the ballots who got contributions by NOT EMPLOYED people.
# when increasing the max_size also here it becomes even more clear how great the difference between Sanders, Bernard and the other candidates is in this aspect.
# NOT EMPLOYED probably also includes college students! Which makes sense because Sanders wants to take away college debt, so he has a big bunch of the students on his side (I heard of 83% somewhere)

# let's look at this in another way:
ggplot(subset(df, contb_emp_status == 'not employed'), aes(contb_emp_status, fill=cand_nm)) + scale_x_discrete() +
  geom_bar()
# time to make a pie chart :)


# wondering which party do IT people lean towards. :)
# collecting all the unique jobs present in the data.frame
jobs <- unique(df$contbr_occupation)
jobs
# haha, found one 'GRANDPA !!' :)
# well, these are too many, and people just put what they wanted. it's not been scanned through and grouped it seems.
# I don't want to get into this.
names(df)

```

# Discussion of the Analysis

### Structure of the dataset

The dataset consists mainly of categorical variables, such as `cand_nm`, `contbr_occupation` or `contbr_zip`. All these are interesting to put in context with the one continuous variable `cont_receipt_amt`, and of course the **amount of rows**.

A more thorough overview of the data can be taken from the official site:
[ftp://ftp.fec.gov/FEC/Presidential_Map/2016/DATA_DICTIONARIES/CONTRIBUTOR_FORMAT.txt](ftp://ftp.fec.gov/FEC/Presidential_Map/2016/DATA_DICTIONARIES/CONTRIBUTOR_FORMAT.txt)

### Main feature(s) of interest in the dataset

How is the number of individual contributions and the amount of money gained through them spread out among the different candidates.
In further abstraction it is interesting to look at the same with aggregating or subsetting the dataset (e.g. by gender of the candidate, or by origin city of the contributor).

### Other features that help to support the investigation into the features of interest

I found it interesting to add the population count of the cities in CA, to be able to calculate a ratio regarding how many contributions and how much money in total, respectively, was given by city in relations to the number of inhabitants.

### New variables created (indirectly) from existing variables

I created three new columns containing some new variables:

- **cand_gender**, defining the gender of the presidential candidate (`'F'` for "female" and `'M'` for "male")
- **cand_party**, defining the party affiliation of the candidate (`'D'` for "Democrats", `'R'` for "Republicans", and `'G'` for "Greens")
- **POPESTIMATE2012**, containing the population estimates of the CA cities for 2012, taken from a different governmental source online
- **contb_emp_status**, where I reduced the `contbr_occupation` to show only whether a contributor was employed or not

### **Investigated features**: unusual distributions? Did you perform any operations on the data to tidy, adjust, or change the form of the data? If so, why did you do this?

---

### Relationships observed in the bivariate part of the investigation. Variations of the investigated features in relation with other features of the dataset.

### Interesting relationships between other features (not the main feature(s) of interest)

### Strongest relationship discovered

In my opinion the strongest relationship that I kept discovering over different plots, was the amount of contributions received, that was very clearly "won" by Sanders, Bernard.

---

### Relationships observed in the multivariate part of the investigation. Features that strengthened each other in terms of looking at your feature(s) of interest

### Interesting or surprising interactions between features

### OPTIONAL: Models created with the dataset. Strengths and limitations of the model.

------

# Final Plots and Summary

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection

# References:

**R:**

- http://stackoverflow.com/questions/8854046/duplicate-row-names-are-not-allowed-error-in-r-programming
- http://stackoverflow.com/questions/4605206/drop-data-frame-columns-by-name
- http://stackoverflow.com/questions/28195996/count-number-of-rows-matching-a-criteria
- http://stackoverflow.com/questions/2626567/collapsing-data-frame-by-selecting-one-row-per-group
- http://stackoverflow.com/questions/4605206/drop-columns-in-r-data-frame
- https://www.miskatonic.org/2012/09/24/counting-and-aggregating-r/
- http://www.r-bloggers.com/r-function-of-the-day-cut/
- http://stackoverflow.com/questions/13093931/r-remove-last-word-from-string
- https://stat.ethz.ch/R-manual/R-devel/library/base/html/merge.html
- http://stackoverflow.com/questions/16516593/convert-from-lowercase-to-uppercase-all-values-in-all-character-variables-in-dat
- http://stackoverflow.com/questions/13967063/remove-duplicate-rows-in-r

**ggplot:**

- http://stackoverflow.com/questions/34963138/error-stat-count-in-ggplot2
- http://stackoverflow.com/questions/3744178/ggplot2-sorting-a-plot

**US politics:**

- https://en.wikipedia.org/wiki/Republican_Party_presidential_primaries,_2016
- https://en.wikipedia.org/wiki/Democratic_Party_presidential_primaries,_2016
- https://en.wikipedia.org/wiki/Jill_Stein
- https://ballotpedia.org/Campaign_finance_requirements_in_California
- https://ballotpedia.org/California_Proposition_34,_Limits_on_Campaign_Contributions_(2000)